Psychological Needs During Intergroup Contact: Three Experience Sampling Studies

### 2022-11-04 Robustness-Methods-and-Results.Rmd
robustness appendix added

Based on our main hypotheses, the aim of our first study was to specifically test the general contact hypothesis, the influence of core need fulfillment, and perceived interaction quality during intergroup contacts. To this aim, we conducted an intensive longitudinal survey study with recent migrants to the Netherlands. Data was collected from May 5\textsuperscript{th} through June 6\textsuperscript{th}, 2018 (and all participants started the study within the first two days).

@@

# look at frequencies of characteristics 
table(workerSampleInfo$gender)
table(workerSampleInfo$edu)
table(workerSampleInfo$occupation)
table(as.character(workerSampleInfo$nationality))

# categorical variables:
# - outgroup contact
# - non-outgroup contact

# continuous variables:
# - Core Need fulfillment
# - SDT Competence
# - SDT Autonomy
# - SDT Relatedness
# - Interaction Quality
# - Outgroup Attitudes
######
# within cor
# between cor
# ---
# Grand mean
# SD between
# SD within
# Proportion of variance
# __ __ __
# Note:
# N within ppt.s 
######

@@

```{r workerVarDescr, include=FALSE}
workerMlCor <-
  MlCorMat(
    data = dtWorkerSupp$workerInteractionType,
    id = "PID",
    selection = c("keyMotiveFulfilled", "competence.daily.all", "autonomy.daily.all", "relatedness.daily.all", "quality_overall_1", "thermometerDutch_1"),
    labels = c("Core Need", "Competence", "Autonomy", "Relatedness", "Quality", "Attitudes NL")
  )

workerMlCor %>%
  kable(
    .,
    caption = "Worker: Multilevel Core Variable Descriptives",
    format = "latex",
    booktabs = TRUE,
    linesep = linesep(c(ncol(.))),
    align = c("l", rep("c", ncol(.) - 1))
  ) %>%
  footnote(
    general = c(
      "Upper triangle: Between-person correlations;",
      "Lower triangle: Within-person correlations;",
      "*** p < .001, ** p < .01,  * p < .05"
    )
  ) %>%
  kable_styling(latex_options = "scale_down") %>%
  gsub("\\begin{table}", "\\begin{table}\n\\begin{minipage}[t][\\textheight][t]{\\textwidth}", ., fixed = TRUE) %>% # fix table position
  gsub("\\end{table}", "\\end{minipage}\n\\end{table}", ., fixed = TRUE) %>% # fix table position
  save_kable("Tables/workerVarDescr.tex")
```

```{r workerOutVarDescr, include=FALSE}
workerOutMlCor <-
  MlCorMat(
    data = dtWorkerSupp$workerInteractionType %>% filter(OutgroupInteraction == "Yes"),
    id = "PID",
    selection = c("keymotive_fulfillemt_1", "competence_1", "autonomy_1", "relatedness_1", "quality_overall_1", "thermometerDutch_1"),
    labels = c("Core Need", "Competence", "Autonomy", "Relatedness", "Quality", "Attitudes NL")
  )

workerOutMlCor %>%
  kable(
    .,
    caption = "Worker: Multilevel Core Variable Descriptives (Outgroup Contact Only)",
    format = "latex",
    booktabs = TRUE,
    linesep = linesep(c(ncol(.))),
    align = c("l", rep("c", ncol(.) - 1))
  ) %>%
  footnote(
    general = c(
      "Upper triangle: Between-person correlations;",
      "Lower triangle: Within-person correlations;",
      "*** p < .001, ** p < .01,  * p < .05"
    )
  ) %>%
  kable_styling(latex_options = "scale_down") %>%
  gsub("\\begin{table}", "\\begin{table}\n\\begin{minipage}[t][\\textheight][t]{\\textwidth}", ., fixed = TRUE) %>% # fix table position
  gsub("\\end{table}", "\\end{minipage}\n\\end{table}", ., fixed = TRUE) %>% # fix table position
  save_kable("Tables/workerOutVarDescr.tex")

# misty::multilevel.descript(dtWorkerSupp$workerWithinBetween$keyMotiveFulfilled, cluster = dtWorkerSupp$workerWithinBetween$PID)
```

\subsubsection{Participants} After receiving ethical approval from the University Masked for Peer Review, we recruited `r nrow(workerSampleInfo)` migrants using the local paid participant pool and specifically targeted non-Dutch migrants to participate in our study. Participants reported on their interactions for at least 30 days with two daily measures (capturing the morning and afternoon). With this design, we aimed at getting 50-60 measurements per participant (\textit{M} = `r format(round(mean(workerSampleInfo$dailiesN),2), nsmall=2)`, \textit{SD} = `r format(round(sd(workerSampleInfo$dailiesN),2), nsmall=2)`, \textit{total N} = `r format(sum(workerSampleInfo$dailiesN), big.mark=",")`). This is a common number of measurements found in experience sampling studies and offers sufficient power to model processes within and between participants \citep[e.g., for a systematic review see][]{AanhetRot2012}. Participants were compensated for their participation with up to 34 Euros – each two Euros for pre- and post-questionnaire as well as 50 Eurocents for every experience sampling measurement occasion. The sample consisted of relatively young, educated, and western migrants from the global north ($M_{age}$ = `r format(round(mean(workerSampleInfo$age),2), nsmall=2)`, $SD_{age}$ = `r format(round(sd(workerSampleInfo$age),2), nsmall=2)`, `r nrow(workerSampleInfo %>% filter(gender == "Female"))` women, `r nrow(workerSampleInfo %>% filter(occupation == "student"))` students). The sample accurately describes the largest groups of migrants in the region \citep[][]{GemeenteGroningen2015}.

\subsubsection{Procedure} The study itself consisted of three main parts, an introductory pre-measurement, and the daily experience sampling measurements, as well as a concluding post-measurement. After giving informed consent, participants started by filling in an online pre-questionnaire assessing demographics and general information about their immigration. Over the next thirty days, the participants were then invited twice a day (at 12 pm and 7pm) to reflect upon their interactions, psychological need fulfillments, and current attitudes towards the Dutch outgroup (\textit{median survey duration} = `r seconds_to_period(median(dtWorker$clean$Duration__in_seconds_)) %>% as.character`, \textit{MAD duration} = `r seconds_to_period(round(mad(dtWorker$clean$Duration__in_seconds_))) %>% as.character`). General compliance was high (`r format(round(sum(workerSampleInfo$dailiesN)/(nrow(workerSampleInfo)*62)*100,2), nsmall=2)`\% of all invited surveys were filled in)\footnote{Two participants completed only two days (among the others, participation was `r format(round(sum(workerSampleInfo$dailiesN[workerSampleInfo$dailiesN>3])/(nrow(workerSampleInfo[workerSampleInfo$dailiesN>3,])*62)*100,2), nsmall=2)`\%).}. The response rates were approximately equal during mornings (\textit{n} = `r format(sum(workerSampleInfo$morningN), big.mark=",")`) and afternoons (\textit{n} = `r format(sum(workerSampleInfo$afternoonN), big.mark=",")`) and most measurements were completed within four hours of the invitation. After the final day of experience sampling measurements, participants were invited to fill in a longer post measurement survey that mirrored the pre-measurement. All key variables for this study were part of the short experience sampling surveys.

@@

\paragraph{Intergroup Contact}To test the prerequisite effect of intergroup contact, every experience sampling measurement started with the question “\textit{Did you meet a Dutch person this morning [/afternoon]? (In person interaction for at least 10 minutes)}”. Our participants recorded between `r dtWorkerSupp$workerAvFreQual$SumContactNL[dtWorkerSupp$workerAvFreQual$SumContactNL > 0] %>% min`--`r dtWorkerSupp$workerAvFreQual$SumContactNL[dtWorkerSupp$workerAvFreQual$SumContactNL > 0] %>% max`
interactions with Dutch outgroup members (`r dtWorkerSupp$workerAvFreQual$PercContactNL[dtWorkerSupp$workerAvFreQual$PercContactNL > 0] %>% min %>% round(2) %>% format(nsmall=2)`--`r dtWorkerSupp$workerAvFreQual$PercContactNL[dtWorkerSupp$workerAvFreQual$PercContactNL > 0] %>% max %>% round(2) %>% format(nsmall=2)`\% of individual experience sampling measurements; `r format(sum(dtWorkerSupp$workerAvFreQual$SumContactNL), big.mark=",")` of all `r format(nrow(dtWorkerSupp$workerInteractionType), big.mark=",")` experience sampling responses)\footnote{Two participants only recorded two experience sampling measurements each and non of these included outgroup contacts. These participants are removed from any analyses that focus on outgroup contacts.}.

\paragraph{Psychological Needs}Irrespective of whether participants had an interaction with Dutch people or not, everyone answered a short series of questions on psychological need fulfillment. However, whereas participants with interactions reported on the need fulfillment during the interaction, people without interactions with Dutch people judged the past daytime period in general. To assess the fulfillment of psychological needs, we included two types of need measurement: (1) the core situational need and (2) general self-determination theory needs.

For the core situational need, we asked participants in an open ended text field: “\textit{What was your most important goal [during the interaction / this morning / this afternoon]?}”. Then, with reference to the text entry, we asked how much this core need was fulfilled during the interaction or the past daytime period: “\textit{[The interaction / You] fulfilled your goal: [-previous text entry-]}” on a continuous slider scale ranging from strongly disagree (-50) to strongly agree (+50).

We, additionally, included a common measure of three self-determination theory needs \citep[see][]{Downie2008}. The items were introduced either by “\textit{During the interaction:}” or “\textit{This morning [/afternoon]:}” and measured autonomy (“\textit{I was myself.}”), competence (“\textit{I felt competent.}”), and relatedness (without intergroup contact “\textit{I had a strong need to belong}”; with intergroup contact: “\textit{I shared information about myself.}” and “\textit{The other(s) shared information about themselves.}”). All items were rated on a continuous slider scale from very little (-50) to a great deal (+50).

\paragraph{Perceived Interaction Quality}As an explanatory mechanism, we assessed ratings of the perceived interaction quality. As our main measurement, participants rated the statement “\textit{Overall the interaction was …}” on two continuous slider scales measuring pleasantness \citep[from unpleasant (-50) to pleasant (+50)) and meaningfulness (from superficial (-50) to meaningful (+50); both items adapted from][]{Downie2008}.

\paragraph{Outgroup Attitudes}At the end of every experience sampling measurement we asked all participants about their current attitudes towards the Dutch – our main dependent variable. To assess the momentary outgroup evaluation we used the common feeling thermometer: “How favorable do you feel towards the Dutch?” \citep[][]{Lavrakas2008}. Participants then rated their attitude on a continuous slider scale from “very cold – 0” through “no feeling – 50” to “very warm – 100”. Both the question phrasing as well as the tick labels were consistent with large-scale panel surveys \citep[e.g.,][]{DeBell2010}.

@@

```{r WorkerFreqAttCor, include=FALSE}
dtWorkerSupp$workerAvFreQual <-
  dtWorkerSupp$workerAvFreQual %>%
  mutate(
    SumContactNL_c = SumContactNL - mean(SumContactNL, na.rm = TRUE),
    SumContactNLAll_c = SumContactNLAll - mean(SumContactNLAll, na.rm = TRUE),
    AvAttitude_c = AvAttitude - mean(AvAttitude, na.rm = TRUE),
    AvQuality_c = AvQuality - mean(AvQuality, na.rm = TRUE)
  )

# correlation panel
pairs.panels.new(
  dtWorkerSupp$workerAvFreQual %>% select(SumContactNL, SumContactNLAll, AvQuality, AvAttitude),
  labels = c(
    "Sum:\nNumer of beeps with Outgroup Contact (NL)",
    "Sum:\nNumber of Outgroup Contacts (NL)",
    "Mean:\nInteraction Quality",
    "Mean:\nOutgroup Attitudes (NL)"
  )
)

# correlation panel with interaction sums winsorized
pairs.panels.new(
  dtWorkerSupp$workerAvFreQual %>% select(WinSumContactNL, WinSumContactNLAll, AvQuality, AvAttitude),
  labels = c(
    "Sum:\nNumer of beeps with Outgroup Contact (NL)\n[Winsorized]",
    "Sum:\nNumber of Outgroup Contacts (NL)\n[Winsorized]",
    "Mean:\nInteraction Quality",
    "Mean:\nOutgroup Attitudes (NL)"
  )
)

rContactAvg <- dtWorkerSupp$workerAvFreQual %>% 
  select(SumContactNL, SumContactNLAll, AvAttitude) %>%
  as.matrix %>%
  rcorr

rContactAvgMax <- format(round(max(abs(rContactAvg$r[rownames(rContactAvg$r) != "AvAttitude","AvAttitude"])),2), nsmall = 2)
pContactAvgMin <- format(round(min(rContactAvg$P[rownames(rContactAvg$P) != "AvAttitude","AvAttitude"]),3), nsmall = 3)

# Average Interaction Quality and their Average Outgroup Attitudes
rAttitudeQuality <- dtWorkerSupp$workerAvFreQual %>% 
  select(AvQuality, AvAttitude) %>%
  as.matrix %>%
  rcorr
```

```{r workerModelOlsAttFreqQual, include=FALSE}
# create list to store Worker models
mdlWorker <- list()

# regression
mdlWorker$lmAttFreqQualX <-
  lm(AvAttitude ~ SumContactNL_c * AvQuality_c, data = dtWorkerSupp$workerAvFreQual)
# summary(lmWorkerAttFreqQualX)

summ(
  mdlWorker$lmAttFreqQualX,
  confint = TRUE,
  digits = 3
)

mdlWorker$lmAttFreqQualXEta <-
  effectsize::eta_squared(mdlWorker$lmAttFreqQualX, partial = TRUE)

interactions::interact_plot(
  mdlWorker$lmAttFreqQualX,
  pred = AvQuality_c,
  modx = SumContactNL_c,
  interval = TRUE,
  partial.residuals = TRUE
)

interactions::johnson_neyman(mdlWorker$lmAttFreqQualX,
                             pred = AvQuality_c,
                             modx = SumContactNL_c,
                             alpha = .05)

@@

# Save variances
mdlWorker$varAttNull <- 
  VarCorr(mdlWorker$lmeAttNull) # save variances
# The estimate of (between-group or Intercept variance, tau_{00}^2):
mdlWorker$tauAttNull <- 
  as.numeric(mdlWorker$varAttNull[1])
# and the estimate of (within-group or residual variance, sigma^2) is:
mdlWorker$sigmaAttNull <- 
  as.numeric(mdlWorker$varAttNull[2])
# The ICC estimate (between/between+within) is:
mdlWorker$IccAttNull <-
  (as.numeric(mdlWorker$varAttNull[1]) / (as.numeric(mdlWorker$varAttNull[1]) + as.numeric(mdlWorker$varAttNull[2])))
mdlWorker$IccPercAttNull <-
  ((as.numeric(mdlWorker$varAttNull[1]) / (as.numeric(mdlWorker$varAttNull[1]) + as.numeric(mdlWorker$varAttNull[2])))) * 100
```
```{r workerModelInterceptAttType, include=FALSE}
# Create and save Model
mdlWorker$lmeInterceptAttType <-
  lme(
    thermometerDutch_1 ~ OutgroupInteractionC + NonOutgroupInteractionC + OutgroupInteractionM + NonOutgroupInteractionM,
    random =  ~ 1 | PID,
    na.action=na.omit,
    data = dtWorkerSupp$workerInteractionType
  )

# Get summary with p-values (Satterthwaite's method)
summ(
  mdlWorker$lmerInterceptAttType <- lmer(
    thermometerDutch_1 ~ OutgroupInteractionC + NonOutgroupInteractionC + OutgroupInteractionM + NonOutgroupInteractionM + (1 | PID),
    data = dtWorkerSupp$workerInteractionType
  ),
  confint = TRUE,
  digits = 3
)

mdlWorker$lmerInterceptAttTypeCI <- 
  confint(method = "Wald", mdlWorker$lmerInterceptAttType)

# Compare new model to previous step
anova(mdlWorker$lmeAttNull, 
      mdlWorker$lmeInterceptAttType) %>%
  as.data.frame() %>%
  select(-call) %>%
  mutate(
    L.Ratio = round(L.Ratio, 3),
    `p-value` = ifelse(`p-value`>=.001, round(`p-value`, 3), "< .001")
  ) %>%
  replace(is.na(.), "") %>%
  add_rownames(., var = "Description") %>%
  mutate(Description = gsub(".*\\$", "", Description)) %>%
  kbl(
    .,
    caption = "Worker: Model Comparison",
    format = "latex",
    linesep = "",
    booktabs = T,
    align = rep("c", ncol(.)),
    digits = 3
  ) %>%
  kable_styling(position = "left")

# Save variances
mdlWorker$varInterceptAttType <- 
  lme4::VarCorr(mdlWorker$lmeInterceptAttType)
```

\subsubsection{Contact Hypothesis}
<!-- We tested the most general contact hypothesis in two steps. First, we assessed whether overall between participants more intergroup interactions were related to to more positive average outgroup attitudes. Second, we tested whether a potential positive effect on outgroup attitudes depended on the interaction quality (jointly with the number of interactions). We find that neither the overall number of contacts nor the number of experience sampling responses with a contact were significantly related with the average outgroup attitudes (all | \textit{r} | < `r rContactAvgMax`, all \textit{p} > `r pContactAvgMin`). However, we did find a medium sized correlation between the participants' average contact quality and their average outgroup attitudes (\textit{r} = `r format(round(rAttitudeQuality$r[rownames(rAttitudeQuality$r) != "AvAttitude","AvAttitude"],2), nsmall = 2)`, \textit{p} = `r format(round(rAttitudeQuality$P[rownames(rAttitudeQuality$r) != "AvAttitude","AvAttitude"],3), nsmall = 3)`). And when considering the number of contacts and average contact quality jointly in a linear regression, we additionally found a statistically significant interaction term (\textit{b} = `r coef(summary(mdlWorker$lmAttFreqQualX))["SumContactNL_c:AvQuality_c","Estimate"] %>% round(2) %>% format(nsmall=2)`, \textit{t}(`r mdlWorker$lmAttFreqQualX$df.residual`) = `r coef(summary(mdlWorker$lmAttFreqQualX))["SumContactNL_c:AvQuality_c","t value"] %>% round(2) %>% format(nsmall=2)`, \textit{p} = `r coef(summary(mdlWorker$lmAttFreqQualX))["SumContactNL_c:AvQuality_c","Pr(>|t|)"] %>% round(3) %>% format(nsmall=3)`, $\eta_p^2$ = `r mdlWorker$lmAttFreqQualXEta %>% filter(Parameter == "SumContactNL_c:AvQuality_c") %>% select(Eta2_partial) %>% as.numeric %>% round(2) %>% format(nsmall=2)`), where in our sample with an increasing number of contacts the positive effect of average contact quality becomes weaker (also see Table \ref{tab:intergroupGeneralTblLong}).  -->
We used a multilevel regression to check whether having an interaction with an outgroup member had a situational (i.e., contemporaneous) effect within the participants. We find that having an outgroup contact is indeed associated with significantly more positive outgroup attitudes (`r MlCoeffLatex(lmeMdl = mdlWorker$lmeInterceptAttType, lmerCI = mdlWorker$lmerInterceptAttTypeCI, varName = "OutgroupInteractionC")`), even after controlling for having an interaction with a non-Dutch (which did not relate to outgroup attitudes independently). Additionally, while multi-level regressions are generally robust against unequal cell sizes, we correct for inequalities by use centered predictors and reintroducing the means as level two predictors. Nonetheless, we also offer an analysis of the joint data sets in Online Supplemental Material A (for full results see Table \ref{tab:intergroupGeneralTblLong}, Figure \ref{fig:ContactHypothesis}, and Online Supplementary Material A)\footnote{Interestingly, adding random slopes to this model did not explain additional variance. This is unusual and might indicate that the effect is very consistent across participants. However, the small number of participants, or other measurement issues provide an alternative explanation, which is why we offer a combined data set analyses following the individual studies.}. 
Thus, in our first data we find initial evidence that outgroup contacts show a positive effect on outgroup attitudes within real-life data. While this offers support for the most basic premise of the intergroup contact hypothesis (i.e., that any interaction will lead to more positive outgroup attitudes), this analysis does not consider any notion of interaction quality. We assess the conditional effects of interaction numbers and interaction quality ratings after we report on the individual studies to avoid under-powered analyses.

\subsubsection{Core Need}The main proposal of our article is that the success of an outgroup contact might be explained by whether or not the contact fulfilled the person's core situational need. This should, in turn, be due to a higher perceived contact quality. We sequentially test whether the fulfillment of the core need during an interaction is (1) related to more positive outgroup attitudes, (2) higher perceived contact quality, and (3) whether the variance explained by the core need is subsumed by the perceived contact quality if considered jointly (for full regression results see Table \ref{tab:intergroupNeedsTblLong} and Figure \ref{fig:AllportNeedFulfillment}).

@@

# Save variances
mdlWorkerOut$Att$varNull <- 
  VarCorr(mdlWorkerOut$Att$lmeNull) # save variances
# The estimate of (between-group or Intercept variance, tau_{00}^2):
mdlWorkerOut$Att$tauNull <- 
  as.numeric(mdlWorkerOut$Att$varNull[1])
# and the estimate of (within-group or residual variance, sigma^2) is:
mdlWorkerOut$Att$sigmaNull <- 
  as.numeric(mdlWorkerOut$Att$varNull[2])
# The ICC estimate (between/between+within) is:
mdlWorkerOut$Att$IccNull <-
  (as.numeric(mdlWorkerOut$Att$varNull[1]) / (as.numeric(mdlWorkerOut$Att$varNull[1]) + as.numeric(mdlWorkerOut$Att$varNull[2])))
mdlWorkerOut$Att$IccPercNull <-
  ((as.numeric(mdlWorkerOut$Att$varNull[1]) / (as.numeric(mdlWorkerOut$Att$varNull[1]) + as.numeric(mdlWorkerOut$Att$varNull[2])))) * 100
```
```{r workerOutModelInterceptAttCore, include=FALSE}
# Create and save Model
mdlWorkerOut$Att$lmeInterceptCore <-
  lme(
    thermometerDutch_1 ~ keymotive_fulfillemt_1_cwc,
    random = ~ 1 | PID,
    na.action=na.omit,
    data = dtWorkerSupp$workerOutWithinBetween
  )

# Get summary with p-values (Satterthwaite's method)
summ(
  mdlWorkerOut$Att$lmerInterceptCore <-
    lmer(
      thermometerDutch_1 ~ keymotive_fulfillemt_1_cwc + (1 | PID),
      data = dtWorkerSupp$workerOutWithinBetween
    ),
  confint = TRUE,
  digits = 3,
  center = FALSE
)

mdlWorkerOut$Att$lmerInterceptCoreCI <- 
  confint(method = "Wald", mdlWorkerOut$Att$lmerInterceptCore)

# Compare new model to previous step
anova(mdlWorkerOut$Att$lmeNull, 
      mdlWorkerOut$Att$lmeInterceptCore) %>%
  as.data.frame() %>%
  select(-call) %>%
  mutate(
    L.Ratio = round(L.Ratio, 3),
    `p-value` = ifelse(`p-value`>=.001, round(`p-value`, 3), "< .001")
  ) %>%
  replace(is.na(.), "") %>%
  kbl(
    .,
    caption = "Worker: Model Comparison",
    format = "latex",
    linesep = "",
    booktabs = T,
    align = rep("c", ncol(.)),
    digits = 3
  ) %>%
  kable_styling(position = "left")

# Save variances
mdlWorkerOut$Att$varInterceptCore <-
  lme4::VarCorr(mdlWorkerOut$Att$lmeInterceptCore)

# The estimate of between-group (or Intercept variance) explained:
# Variance Explained = 1 – (Var with Predictor/Var without Predictor)
mdlWorkerOut$Att$varBtwInterceptCore <-
  1 - (mdlWorkerOut$Att$varInterceptCore["(Intercept)", "Variance"] %>% as.numeric / mdlWorkerOut$Att$varNull["(Intercept)", "Variance"] %>% as.numeric)
mdlWorkerOut$Att$varBtwPercInterceptCore <-
  (1 - (mdlWorkerOut$Att$varInterceptCore["(Intercept)", "Variance"] %>% as.numeric / mdlWorkerOut$Att$varNull["(Intercept)", "Variance"] %>% as.numeric)) * 100
# and the estimate of within-group (or residual variance) explained is:
mdlWorkerOut$Att$varWithinInterceptCore <-
  1 - (mdlWorkerOut$Att$varInterceptCore["Residual", "Variance"] %>% as.numeric / mdlWorkerOut$Att$varNull["Residual", "Variance"] %>% as.numeric)
mdlWorkerOut$Att$varWithinPercInterceptCore <-
  (1 - (mdlWorkerOut$Att$varInterceptCore["Residual", "Variance"] %>% as.numeric / mdlWorkerOut$Att$varNull["Residual", "Variance"] %>% as.numeric)) * 100
```
```{r workerOutModelSlopesAttCore, include=FALSE}
# Create and save Model (optimizer needed to reach convergence)
mdlWorkerOut$Att$lmeSlopesCore <-
  lme(
    thermometerDutch_1 ~
      keymotive_fulfillemt_1_cwc,
    random = ~ 1 + keymotive_fulfillemt_1_cwc | PID,
    control = lmeControl(opt = "optim"),
    na.action=na.omit,
    data = dtWorkerSupp$workerOutWithinBetween
  )

# Get summary with p-values (Satterthwaite's method) [+ save model for plotting]
summ(
  mdlWorkerOut$Att$lmerSlopesCore <- lmer(
    thermometerDutch_1 ~
      keymotive_fulfillemt_1_cwc +
      (1 + keymotive_fulfillemt_1_cwc | PID),
    data = dtWorkerSupp$workerOutWithinBetween
  ),
  confint = TRUE,
  digits = 3,
  center = FALSE
)

# all variables standardized within PPT
summ(
  mdlWorkerOut$Att$lmerSlopesCoreZ <- lmer(
    thermometerDutch_1_zwc ~
      keymotive_fulfillemt_1_zwc +
      (1 + keymotive_fulfillemt_1_zwc | PID),
    data = dtWorkerSupp$workerOutWithinBetween
  ),
  confint = TRUE,
  digits = 3,
  center = FALSE
)

# standardized coefficients
stdCoef.merMod(mdlWorkerOut$Att$lmerSlopesCore)

# 95%CIs
mdlWorkerOut$Att$lmerSlopesCoreCI <- 
  confint(method = "Wald", mdlWorkerOut$Att$lmerSlopesCore)

# Attempts at R^2
r2mlm::r2mlm(mdlWorkerOut$Att$lmerSlopesCore, bargraph = TRUE)
mitml::multilevelR2(mdlWorkerOut$Att$lmerSlopesCore)
performance::r2(mdlWorkerOut$Att$lmerSlopesCore)
performance::model_performance(mdlWorkerOut$Att$lmerSlopesCore)
performance::compare_performance(mdlWorkerOut$Att$lmerNull, 
                                 mdlWorkerOut$Att$lmerInterceptCore, 
                                 mdlWorkerOut$Att$lmerSlopesCore)

# Compare new model to previous step
anova(mdlWorkerOut$Att$lmeNull, 
      mdlWorkerOut$Att$lmeInterceptCore, 
      mdlWorkerOut$Att$lmeSlopesCore) %>%
  as.data.frame() %>%
  select(-call) %>%
  mutate(
    L.Ratio = round(L.Ratio, 3),
    `p-value` = ifelse(`p-value`>=.001, round(`p-value`, 3), "< .001")
  ) %>%
  replace(is.na(.), "") %>%
  kbl(
    .,
    caption = "Worker: Model Comparison",
    format = "latex",
    linesep = "",
    booktabs = T,
    align = rep("c", ncol(.)),
    digits = 3
  ) %>%
  kable_styling(position = "left")

# Save variances
mdlWorkerOut$Att$varSlopesCore <- 
  lme4::VarCorr(mdlWorkerOut$Att$lmeSlopesCore)

# The estimate of between-group (or Intercept variance) explained:
# Variance Explained = 1 – (Var with Predictor/Var without Predictor)
mdlWorkerOut$Att$varBtwSlopesCore <-
  1 - (mdlWorkerOut$Att$varSlopesCore["(Intercept)", "Variance"] %>% as.numeric / mdlWorkerOut$Att$varInterceptCore["(Intercept)", "Variance"] %>% as.numeric)
mdlWorkerOut$Att$varBtwPercSlopesCore <-
  (1 - (mdlWorkerOut$Att$varSlopesCore["(Intercept)", "Variance"] %>% as.numeric / mdlWorkerOut$Att$varInterceptCore["(Intercept)", "Variance"] %>% as.numeric)) * 100
# and the estimate of within-group (or residual variance) explained is:
mdlWorkerOut$Att$varWithinSlopesCore <-
  1 - (mdlWorkerOut$Att$varSlopesCore["Residual", "Variance"] %>% as.numeric / mdlWorkerOut$Att$varInterceptCore["Residual", "Variance"] %>% as.numeric)
mdlWorkerOut$Att$varWithinPercSlopesCore <-
  (1 - (mdlWorkerOut$Att$varSlopesCore["Residual", "Variance"] %>% as.numeric / mdlWorkerOut$Att$varInterceptCore["Residual", "Variance"] %>% as.numeric)) * 100

# Assumption Checks:
mdlWorkerOut$Att$diagSlopesCore <- 
  sjPlot::plot_model(mdlWorkerOut$Att$lmerSlopesCore, type = "diag")
grid.arrange(
  mdlWorkerOut$Att$diagSlopesCore[[1]],
  mdlWorkerOut$Att$diagSlopesCore[[2]]$`PID`,
  mdlWorkerOut$Att$diagSlopesCore[[3]],
  mdlWorkerOut$Att$diagSlopesCore[[4]]
)

# Plot prediction model
mdlWorkerOut$Att$predSlopesCore <- 
  dtWorkerSupp$workerOutWithinBetween %>%
  select(thermometerDutch_1, session, PID) %>% 
  mutate(measure = predict(mdlWorkerOut$Att$lmeSlopesCore,
                           dtWorkerSupp$workerOutWithinBetween,
                           re.form = NA
                           )
         )

(
  mdlWorkerOut$Att$predPltSlopesCore <-
    ggplot(data = mdlWorkerOut$Att$predSlopesCore, aes(x = session, y = measure)) +
    geom_line(alpha = 1, color = "blue") +
    geom_line(aes(y = thermometerDutch_1), alpha = 1) +
    facet_wrap( ~ PID, ncol = 6) +
    xlab("Time") +
    ylab("Outgroup Attitudes") +
    theme_Publication()
)
ggsave(
  filename = "Figures/WorkerOut_PredictionPlot_SlopesAttCore.png",
  mdlWorkerOut$Att$predPltSlopesCore,
  width = 18,
  height = 12,
  dpi = 800,
  units = "cm",
  device = "png"
)
```

@@

The aim of Study 2 is similar to Study 1, as we again test the general contact hypothesis, the influence of core need fulfillment, and perceived contact quality during intergroup contacts. However, in this second study we collected a substantially larger sample of international students who recently arrived in the Netherlands and also improved the study design (e.g., pop-up explanations described later). The survey method again offers a large body of ecologically valid data on need satisfaction in real-life intergroup contact situations as these students will likely interact with the Dutch majority outgroup on a daily basis. Data was collected from November 19\textsuperscript{th}, 2018 through January 6\textsuperscript{th}, 2019.

@@

# categorical variables:
# - outgroup contact
# - non-outgroup contact

# continuous variables:
# - Core Need fulfillment
# - SDT Competence
# - SDT Autonomy
# - SDT Relatedness
# - Interaction Quality
# - Outgroup Attitudes
######
# within cor
# between cor
# ---
# Grand mean
# SD between
# SD within
# Proportion of variance
# __ __ __
# Note:
# N within ppt.s 
######

@@

```{r studentVarDescr, include=FALSE}
studentMlCor <-
  MlCorMat(
    data = dtStudentSupp$studentInteractionType,
    id = "PID",
    selection = c("KeyNeedFullfillment", "Competence", "Autonomy", "Relatedness", "quality_overall", "AttitudesDutch"),
    labels = c("Core Need", "Competence", "Autonomy", "Relatedness", "Quality", "Attitudes NL")
  )

studentMlCor %>%
  kable(
    .,
    caption = "Student: Multilevel Core Variable Descriptives",
    format = "latex",
    booktabs = TRUE,
    linesep = linesep(c(ncol(.))),
    align = c("l", rep("c", ncol(.) - 1))
  ) %>%
  footnote(
    general = c(
      "Upper triangle: Between-person correlations;",
      "Lower triangle: Within-person correlations;",
      "*** p < .001, ** p < .01,  * p < .05"
    )
  ) %>%
  kable_styling(latex_options = "scale_down") %>%
  gsub("\\begin{table}", "\\begin{table}\n\\begin{minipage}[t][\\textheight][t]{\\textwidth}", ., fixed = TRUE) %>% # fix table position
  gsub("\\end{table}", "\\end{minipage}\n\\end{table}", ., fixed = TRUE) %>% # fix table position
  save_kable("Tables/studentVarDescr.tex")
```

```{r studentOutVarDescr, include=FALSE}
studentOutMlCor <-
  MlCorMat(
    data = dtStudentSupp$studentInteractionType %>% filter(OutgroupInteraction == "Yes"),
    id = "PID",
    selection = c("KeyNeedFullfillment", "Competence", "Autonomy", "Relatedness", "quality_overall", "AttitudesDutch"),
    labels = c("Core Need", "Competence", "Autonomy", "Relatedness", "Quality", "Attitudes NL")
  )

>>>>

@@

@@

@@

\paragraph{General Psychological Needs} In addition to the intergroup contact dummy and situational core need reported in the main text, we included a common measure of three self-determination theory needs \citep[see][]{Downie2008}. The items were introduced either by “\textit{During the interaction:}” or “\textit{This morning [/afternoon]:}” and measured autonomy (“\textit{I was myself.}”), competence (“\textit{I felt competent.}”), and relatedness (without intergroup contact “\textit{I had a strong need to belong}”; with intergroup contact: “\textit{I shared information about myself.}” and “\textit{The other(s) shared information about themselves.}”). All items were rated on a continuous slider scale from very little (-50) to a great deal (+50).

@@

```{r workerModelOlsAttFreqQual, include=FALSE}
# create list to store Worker models
mdlWorker <- list()
```

@@

# Save variances
mdlWorker$varAttNull <- 
  VarCorr(mdlWorker$lmeAttNull) # save variances
# The estimate of (between-group or Intercept variance, tau_{00}^2):
mdlWorker$tauAttNull <- 
  as.numeric(mdlWorker$varAttNull[1])
# and the estimate of (within-group or residual variance, sigma^2) is:
mdlWorker$sigmaAttNull <- 
  as.numeric(mdlWorker$varAttNull[2])
# The ICC estimate (between/between+within) is:
mdlWorker$IccAttNull <-
  (as.numeric(mdlWorker$varAttNull[1]) / (as.numeric(mdlWorker$varAttNull[1]) + as.numeric(mdlWorker$varAttNull[2])))
mdlWorker$IccPercAttNull <-
  ((as.numeric(mdlWorker$varAttNull[1]) / (as.numeric(mdlWorker$varAttNull[1]) + as.numeric(mdlWorker$varAttNull[2])))) * 100
```

@@

# Save variances
mdlWorkerOut$Att$varNull <- 
  VarCorr(mdlWorkerOut$Att$lmeNull) # save variances
# The estimate of (between-group or Intercept variance, tau_{00}^2):
mdlWorkerOut$Att$tauNull <- 
  as.numeric(mdlWorkerOut$Att$varNull[1])
# and the estimate of (within-group or residual variance, sigma^2) is:
mdlWorkerOut$Att$sigmaNull <- 
  as.numeric(mdlWorkerOut$Att$varNull[2])
# The ICC estimate (between/between+within) is:
mdlWorkerOut$Att$IccNull <-
  (as.numeric(mdlWorkerOut$Att$varNull[1]) / (as.numeric(mdlWorkerOut$Att$varNull[1]) + as.numeric(mdlWorkerOut$Att$varNull[2])))
mdlWorkerOut$Att$IccPercNull <-
  ((as.numeric(mdlWorkerOut$Att$varNull[1]) / (as.numeric(mdlWorkerOut$Att$varNull[1]) + as.numeric(mdlWorkerOut$Att$varNull[2])))) * 100
```

@@

@@

@@
